\textbf{NBA Party}

### 2017-04-24 init_model.rnw
Added section on replicating analysis for MLB.

<<echo=F,results='hide'>>=

@@

load(file = 'tmp_data/covers_lines.RData')
load(file = 'tmp_data/nmusician_estabs.RData')
load(file = 'tmp_data/population_by_county_year.RData')

musicians <- merge(musicians, pop, by = c('county', 'season'), all.x = T)

@@

@@

<<echo=F>>=
musicians[, county := gsub(' (county)|(division),.*$', '', county, ignore.case = T)]
musicians[, list(county, team, season, nmusicians = round(nmusicians, 1))]
@

@@

To end up with something like

@@

\section{A significant model} We keep things simple, and try to explain the likelihood
of meeting the spread as a function of whether a team visited a city with an active night-life
the day before.

@@

<<echo=2:5>>=
require(e1071)
svm.data <- lines[last.game.loc != team, list(hour = hour(date), last.game.time,
                                              party, season, outcome)] %>% na.omit

m <- svm(x = svm.data[season < 2016, -c('outcome', 'season'), with = F],
         y = svm.data[season < 2016, outcome == 'W'],
         scale = T, type = 'C-classification', probability = T, kernel = 'radial')

preds <- predict(m, newdata = svm.data[season == 2016, -c('outcome', 'season'), with = F])
table(preds, svm.data[season == 2016 , outcome])
@

@@

lm(demeaned.reb.per.min ~ party + lag.chg.pos +
       I(log(travel.dist+1)/nhours.lgame) + nhours.lgame,
   data = data[season < 2017 & last.game.loc != team]) %>%
   summary
@

>>>>

<<echo=F,results='hide',messages=F>>=

@@

load(file = 'tmp_data/covers_lines.RData')
load(file = 'tmp_data/nmusician_estabs.RData')
load(file = 'tmp_data/population_by_county_year.RData')
musicians <- merge(musicians, pop, by = c('county', 'season'), all.x = T)

@@

For results, skip to section \ref{seq: significant model}.

@@

<<echo=F,results='hide'>>=
musicians[, county := gsub(' (county)|(division),.*$', '', county, ignore.case = T)]
@

@@

To end up with a quantity that reflects the total number of changes in possession across both teams within each game.

@@

\section{A significant model} 
\label{seq: significant model}
We keep things simple, and try to explain the likelihood
of meeting the spread as a function of whether a team visited a city with an active night-life
the day before.

@@

<<echo=2:5>>=
require(e1071)
svm.data <- lines[last.game.loc != team, 
                  list(hour = hour(date), last.game.time, party, 
                       loss.season, season, outcome)] %>%
    na.omit

### tune.svm(..., gamma = 2^(-5:5))
m <- svm(x = svm.data[season < 2016, -c('outcome', 'season'), with = F],
              y = svm.data[season < 2016, outcome == 'W'],
              scale = T, type = 'C-classification', probability = T, 
              kernel = 'radial')
preds <- predict(m, newdata = svm.data[season == 2016, -c('outcome', 'season'), with = F])
table(preds, svm.data[season == 2016 , outcome])
@

@@

lm(demeaned.reb.per.min ~ party + lag.chg.pos +
       I(log(travel.dist+1)/nhours.lgame) + nhours.lgame,
   data = data[season < 2017 & last.game.loc != team] %>% na.omit) %>%
   summary
@

\section{Robustness Check: Replicating the analysis for MLB}
We go ahead and repeat this process for major league baseball.
The only difference here is that when we define our party variable,
we implicitly interact it with \texttt{weekend}.

<<echo=-1>>=
load(file = 'tmp_data/covers_lines_mlb.RData')
lines[, party := ifelse(last.game.loc != team & weekend == 1, nmusicians, 0)]

m <- glm(I(team.score > opponent.score) ~ party + I(team == location) + 
             odds + nhours.lgame + weekend + travel.dist,
         data = lines, family = 'binomial')
summary(m)
@

We note that although we don't quite achieve conventional statistical significance,
we're not that far off the mark.

\paragraph{Betting on MLB}
We can also earn a profit making predictions, even after accounting for the cut
taken by the betting house.

<<echo=F>>= 
source('functions.r')

m <- glm(I(team.score > opponent.score) ~ party,
         data = lines[year(date) < 2016], family = 'binomial')
p <- predict(m, newdata = lines[year(date) == 2016], type = 'response')

idx <- which(p < .45)
p1 <- mapply(Predict, prediction = p[idx], vegas = lines[year(date) == 2016, odds][idx],
             MoreArgs=list(threshold = 0.01))
b <- mapply(Bet, our.prediction = p1,
            actual.outcome = lines[year(date) == 2016, team.score > opponent.score][idx],
            odds = lines[year(date) == 2016, odds][idx], SIMPLIFY = T)

sum(b, na.rm = T)

If we were to bet one hundred dollars on each game in 2016,
we would earn a profit of just less than \$400 over the course of 45 games.
